### 2016-08-18 transformation.tex
polishing edits

In the following, we illustrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al. \cite{ecrts15nelissen}, which allows analyzing the worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity. Specifically,  
Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} can be applied under the following conditions:
\begin{enumerate}
	\item the task set contains only one self-suspending task, 
	\item the self-suspending task is the lowest-priority task, 
	\item the scheduling policy is preemptive fixed-priority scheduling, and 
	\item all tasks have constrained deadlines (i.e., $D_i \leq T_i$ for all $\tau_i$).
\end{enumerate}

As an aside, it is interesting to note that, even for the restrictive case above,  Nelissen et al.'s approach \cite{ecrts15nelissen}, which is based on \emph{mixed-integer linear programming} (MILP), provides only an upper bound on the worst-case response time, and not necessarily the \emph{exact} worst-case response time. However, at least conceptually, the exact worst-case response time \emph{can} be obtained based on their analysis by combining it with an exhaustive search. Specifically, Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen} provides a characterization of the worst-case release pattern that yields a maximal response time. Thus, by exploring all  release patterns that satisfy the conditions stated in Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen}, an exact bound can be determined, with very high time complexity. A detailed discussion of this approach can be found in a recent report~\cite{Chen2016b}.

Let us now return to the discussion of the task-set transformation needed to apply Theorem 5 in~\cite{Raj:suspension1991}. For an arbitrary number of tasks $k \geq 2$, 
suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$). Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.

@@

However, deriving an exact bound $W_k^j$ for task $\tau_k$ and for $j=1,2,\ldots,m_k-1$ is not easy: 
even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound on the worst-case response time requires exponential time complexity if $j \geq 2$. In fact, it has recently been shown that the problem of verifying the schedulability of such a task set is coNP-hard in the strong sense~\cite{RTSS2016-suspension}. Therefore, it follows analogously that analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense~\cite{RTSS2016-suspension}.

Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their analyses pertain to unrestricted self-suspensions. However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.

>>>>

In the following, we illustrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al. \cite{ecrts15nelissen}, which allows analyzing the worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity. Specifically,  
Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen}, which is based on \emph{mixed-integer linear programming} (MILP), can be applied under the following conditions:
\begin{enumerate}
	\item the task set contains only one self-suspending task, 
	\item the self-suspending task is the lowest-priority task, 
	\item the scheduling policy is preemptive fixed-priority scheduling, and 
	\item all tasks have constrained deadlines (i.e., $D_i \leq T_i$ for all $\tau_i$).
\end{enumerate}

As an aside, it is interesting to note that, even for the restrictive case above,  Nelissen et al.'s MILP-based analysis~\cite{ecrts15nelissen} provides only an upper bound on the worst-case response time, and not necessarily the \emph{exact} worst-case response time. However, at least conceptually, an exact answer \emph{can} be obtained based on their analysis by combining it with an exhaustive search. Specifically, Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen} provides a characterization of the worst-case release pattern that yields a maximal response time. Thus, by exploring all  release patterns that satisfy the conditions stated in Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen}, an exact bound can be determined (at great computational cost). A detailed discussion of this approach can be found in a recent report~\cite{Chen2016b}.

Let us now return to the discussion of the task-set transformation that is needed before Theorem 5 in~\cite{Raj:suspension1991} can be applied. For an arbitrary number of tasks $k \geq 2$, 
suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$). Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.

@@

However, deriving an exact bound $W_k^j$ for task $\tau_k$ and for $j=1,2,\ldots,m_k-1$ is not easy: 
even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound on the worst-case response time requires exponential time complexity if $j \geq 2$. In fact, it has recently been shown that the problem of verifying the schedulability of such a task set is coNP-hard in the strong sense~\cite{RTSS2016-suspension}. It follows analogously that analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense~\cite{RTSS2016-suspension}.

Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their analyses pertain to unrestricted self-suspensions. However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that their observations transfer to the period enforcement case.

### 2016-08-18 untitled.tex
polishing edits

\section{Introduction} 
When real-time tasks suspend themselves (due to blocking I/O, lock contention, \textit{etc.}), they defer a part of their execution to be processed at a later time. A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}. This penalty, which is maximized when a task defers the completion of one job just until the arrival of the next job, can manifest as response-time increases and thus may lead to deadline misses.

@@

The first self-suspension segment $S_i^0$, prior to the first execution segment, is equivalent to release jitter (i.e., the parameter $J_i$ in Section~\ref{sec:jitter}). However, in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ is assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free). Unless noted otherwise we adopt this convention and assume $S_i^0 = 0$.

>>>>

\section{Introduction} 
When real-time tasks suspend themselves (due to blocking I/O, lock contention, \textit{etc.}), they defer a part of their execution to be processed at a later time. A consequence of such deferred execution is a potential interference penalty for lower-priority tasks~\cite{LSS:87,LSST:91,Ra:90,ABRTW:93,SLS:95,WC16-suspend-DATE,ecrts15nelissen}. This penalty, which is maximized when a task defers the completion of one job just until the release of the next job, can manifest as response-time increases and thus may lead to deadline misses.

@@

The first self-suspension segment $S_i^0$, prior to the first execution segment, is equivalent to release jitter (i.e., the parameter $J_i$ in Section~\ref{sec:jitter}). However, in much of the literature on the segmented self-suspending task model, the segment $S_i^0$ is assumed to be absent (i.e., $S_i^0 = 0$), such that there are only $m_i - 1$ suspension intervals (and jobs arrive jitter-free). We adopt this convention and assume $S_i^0 = 0$ unless noted otherwise.
